Automated Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-01 paper/paper_new.tex
Further editing/refinement of Section 5.

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images, e.g. with low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements, provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata,~\cite{biodiv_info_standards} detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images
to determine if, for instance, a specific fish grows to different lengths
in different habitats.

Since the digital collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible. Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics tools and techniques, we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of 7,247
images drawn from the Illinois Natural History Survey (INHS) Collection.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results are presented, followed by a discussion. The conclusion highlights key findings and identifies next steps.

@@

\section{Methods}
Our process for metadata generation can be broken into three steps: object detection with Facebook's Detectron2 machine learning library (subsequently referred to as \verb|detectron|), image informatics analysis at the pixel level, and calculations on the results of the previous steps to determine higher level metadata properties.
% \todo{Definitely need to explain somewhere in more detail about our current admissibility criteria for images. Not sure where I want to put it yet so just leaving it as a todo}

\begin{figure}[H]
  \centering
  \includegraphics[width=.49\linewidth]{images/teaser1_crop}
  %\hspace{4mm}
  \includegraphics[width=.49\linewidth]{images/teaser2_crop}
  \caption{Initial object detection on specimen images using Detectron2~\cite{wu2019detectron2}}
  \label{fig:teaser}
\end{figure}
\subsection{Detectron}
A prerequisite task to performing any advanced metadata property generation
is finding the specimens (and other relevant objects) within the collection
images. Object detection has been a broadly active field of study in recent
years \cite{zou2019object}, and has resulted in a number of well-tested, purpose-built architectures. We elected to use Facebook AI Research's \verb|detectron| tool \cite{wu2019detectron2}, and specifically its implementation of the Mask R-CNN architecture \cite{he2018mask}, for object detection in our project.
\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format datasets. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given image. For our project, we are currently using it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers.

@@

\begin{table*}[t]
    \centering
    \caption{Metadata properties}
    \label{tab:properties}
    \begin{tabular}{cccp{0.5\linewidth}}
        \toprule
        \textbf{Property} & \textbf{Association} & \textbf{Type} & \textbf{Explanation}\\
        \midrule
        \verb|has_fish| & Overall Image & Boolean & Whether a fish was found in the image.\\
        \verb|fish_count| & Overall Image & Integer & The quantity of fish present.\\
        \verb|has_ruler| & Overall Image & Boolean & Whether a ruler was found in the image.\\
        \verb|ruler_bbox| & Overall Image & 4 Tuple & The bounding box of the ruler (if found).\\
        \verb|scale| & Overall Image & Float & The scale of the image in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|bbox| & Per Fish & 4 Tuple & The top left and bottom right coordinates of the bounding box for a fish.\\
        \verb|background.mean| & Per Fish & Float & The mean intensity of the background within a given fish's bounding box.\\
        \verb|background.std| & Per Fish & Float & The standard deviation of the background within a given fish's bounding box.\\
        \verb|foreground.mean| & Per Fish & Float & The mean intensity of the foreground within a given fish's bounding box.\\
        \verb|foreground.std| & Per Fish & Float & The standard deviation of the foreground within a given fish's bounding box.\\
        \verb|centroid| & Per Fish & 4 Tuple & The centroid of a given fish's bitmask.\\
        \verb|primary_axis| & Per Fish & 2D Vector & The unit length primary axis (eigenvector) for the bitmask of a given fish.\\
        \verb|clock_value| & Per Fish & Integer & Fish's primary axis converted into an integer ``clock value'' between 1 and 12.\\
        \verb|length| & Per Fish & Float & The length of a fish in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|mask| & Per Fish & 2D Matrix & The bitmask of a fish in 0's and 1's.\\
        \verb|pixel_analysis_failed| & Per Fish & Boolean & Whether the pixel analysis process failed for a given fish. If \verb|true|, \verb|detectron|'s mask and bounding box were used for metadata generation.\\
        \verb|score| & Per Fish & Float & The percent confidence score output by \verb|detectron| for a given fish.\\
        \verb|has_eye| & Per Fish & Boolean & Whether an eye was found for a given fish.\\
        \verb|eye_center| & Per Fish & 2 Tuple & The centroid of a fish's eye.\\
        \verb|side| & Per Fish & String & The side (i.e.\ \verb|'left'| or \verb|'right'|) of the fish that is facing the camera (dependent on finding its eye).\\
      \bottomrule
\end{tabular}
\end{table*}

@@

@@

Since we recalculate the mask of the fish, we must recalculate its centroid as well. This can be done via
\begin{equation}
    (\bar{x}, \bar{y}) = (\mathrm{round}(\frac{M_{10}}{M_{00}}), \mathrm{round}(\frac{M_{01}}{M_{00}})),
\end{equation}
where \(M_{00}\) is the pixel area of the fish's blob, \(M_{10}\) is the sum of all the \(x\) values of blob pixels, and \(M_{01}\) is the sum of all the \(y\) values of blob pixels.
\subsubsection{side}
Determining which \verb|side| of the fish is visible is predicated on finding its eye. If an eye is found, the sign of the \(x\) component of the vector from the centroid of the fish to the centroid of the eye specifies which side is up: negative for left and positive for right. This assumes the fish was photographed vertically, which is essentially always the case for all image collections our group has worked on.
\subsubsection{primary\_axis and clock\_value}
The \verb|primary_axis| of a fish can be calculated by taking the covariance of its blob in \(x\) and \(y\), which yields its principle eigenvector.
The eigenvector can be directly assigned to the property. If an eye is
present, we ensure that \verb|primary_axis| points in the direction of the
eye in relation to the fish's centroid.

@@

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results and Discussion}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{leipzig2021biodiversity} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel analysis processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 50 specimens from the set of \(7,244\) were analyzed by
hand for comparison.

Our automated process currently generates 7 of the 22 core metadata properties: \verb|if_fish| (\verb|has_fish|), \verb|fish_number| (\verb|fish_count|), \verb|if_ruler| (\verb|has_ruler|), \verb|specimen_angled| (\verb|clock_value|), \verb|specimen_view| (\verb|side|), \verb|brightness| (derived from \verb|background.mean| and \verb|foreground.mean|), and \verb|if_background_uniform| (\verb|background.std|). In addition, our process also calculates bounding boxes and fish lengths in centimeters.

\subsection{Fish Detection}
All images in the INHS dataset contain exactly one fish. For \(7,209\) of the specimen images, one fish was detected, a 99.5\% correct rate.
For 25 of the images, 2 fish were detected, 3 fish were detected for 3 images, and for 7 images no fish were found. In the case of greater than 1 fish, 9 of the 28 contained tags which overlapped the fish and were themselves labeled as a second fish. Of the remaining 17, \verb|detectron| erroneously labeled the fish as two separate fish objects, or labeled a subsection of the fish a second time. The 7 fish that were not detected were quite small, which is something that is a type of specimen that is currently lacking from the training set. See the left image in Figure~\ref{fig:fish_detect_wrong} for an example of this. Fish that were double labeled were generally quite large and/or dissimilar from the fish found in the training set, such as the right image in Figure~\ref{fig:fish_detect_wrong}.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/none1}
  \includegraphics[width=0.49\linewidth]{images/double1}
  \caption{A fish that was not detected (left) and a fish that was detected twice (right).}
  \label{fig:fish_detect_wrong}
\end{figure}

\subsection{Ruler Detection}
For all but 2 of the \(7,244\) images \verb|detectron| was able to find the ruler, a nearly perfect correct rate.
In 56 of the images, the ruler itself was found but the ``2'' and/or the ``3'' on it was not, so a scale calculation could not be performed, producing a 99.2\% success rate for
the \verb|scale| computation.

Images where one of these three objects were not detected generally had some form of coloration issue. They were either washed out, very dark or yellow in hue. Some of the rulers for which ``2'' and/or ``3'' were not detected were particularly scratched and damaged. Only ``3'' was missed in 45 of the 56 cases, only ``2'' was missed in 2 cases, and both were missed in 9 cases. This may indicate that more training samples for ``3'' are required. Many of the rulers where both numerals were missed were particularly small within the image, which again may be solvable through bolstering the training dataset.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/no_ruler1}
  \includegraphics[width=0.49\linewidth]{images/no_ruler2}
  \caption{The two images where the ruler was not found. The left image exhibits a yellow hue, and the right is quite washed out with poor contrast.}
\end{figure}

@@

@@

There were six cases where an eye was detected but the \verb|side| value
was wrong. Here, a spot on the wrong side of the fish was labeled as the
most likely eye within the bounding box of the fish, i.e.~eye detection 
was incorrect.
There were an additional 17 images for which the automated process generated a result that did not match the manually created data. For these remaining cases the manual data was incorrect, giving the automated system an error rate \(2.8\) times lower than the human error rate.
This result highlights the additional utility of automated methods to
double-check and verify human-generated metadata.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/wrong_side_orig}
  \includegraphics[width=0.49\linewidth]{images/wrong_side1}
  \caption{A fish for which a splotch on its tail fin was labeled the most likely eye.}
\end{figure}

\subsection{Clock Value}
Clock position values were successfully generated for \(6,991\) of the images. Of those, all but 8 were within \(\pm{}1\) of the correct result,
our definition of a correct/acceptable result. The correct rate for this
computation is 96.4\%.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/curved1}
  \caption{An example of a heavily curved specimen.}
  \label{fig:curvedFish}
\end{figure}
Of the specimens for which clock values were generated, 33 did not match the manually created data (within a tolerance of \(\pm{}1\)).
For 25 of those, the human-generated data was incorrect, giving the automated process a \(3.1\) times lower error rate. Of the remaining 8, 2 specimens were quite curved making it difficult to assign a clear angle value, as
seen in Figure \ref{fig:curvedFish}.
The other 6 were values of 3 instead of 9 or vice versa, which resulted from
a mislabeled eye as discussed in the previous section.
%\todo{Why is the following sentence here? What does it have to do with
%Clock Value?}
%A small number of images (such as on the right in Figure~\ref{fig:teaser}) have tags that overlap with the fish. These tags were light enough that they were correctly labeled as background by the pixel analysis.

\subsection{Image Contrast}
\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{images/histo}
  \caption{Histogram for difference in brightness between foreground an background.}
  \label{fig:histo}
\end{figure}

The contrast of an image is an important aspect of the image quality for the BGNN study. Ideally, for INHS images the fish should be well lit and vibrant, and the background should also be as light and white as possible. Figure~\ref{fig:histo} shows the distribution of the difference between the mean intensity of the foreground (i.e.\ the pixels of the fish) and the mean intensity of the background within the fish's bounding box. The overall mean of the distribution is \(144.26\), with a standard deviation of \(15.79\). Images on the low end of this distribution exhibit poor foreground--background contrast, and images on the high end likely contain poorly lit specimens. Examples of each type of image (low, normal, and high contrast) can be seen in Figure~\ref{fig:contrast}. The top left image has a mean intensity of 103 (low), the top right 144 (normal), and the bottom 186 (high).

\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/dark}
  \includegraphics[width=0.49\linewidth]{images/normal}
  \includegraphics[width=0.49\linewidth]{images/bright}
  \caption{Examples of low, normal and high contrast fish images.}
  \label{fig:contrast}
\end{figure}

\subsection{Brightness}
Specimen brightness is one of the 22 hand recorded properties. It is encoded as \verb|dark|, \verb|normal| or \verb|bright|. These values correspond to the measurement of mean foreground intensity by the automated system. The correlation between the two measurements can be seen in Table~\ref{tab:bright} below.
\begin{table}[H]
    \centering
    \caption{Brightness measurements}
    \label{tab:bright}
    \begin{tabular}{cccc}
        \toprule
        \textbf{Class} & \textbf{Mean Intensity} & \textbf{Standard Deviation} & \textbf{Instances}\\
        \midrule
        Dark & 75.15 & 13.49 & 1800\\
        Normal & 93.64 & 14.78 & 5186\\
        Bright & 108.39 & 15.21 & 242\\
        \bottomrule
    \end{tabular}
\end{table}

\textbf{Todo}: I suppose more should be said here.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/011196_pred}
  \includegraphics[width=0.49\linewidth]{images/011196_mask}
  \includegraphics[width=0.49\linewidth]{images/52399_pred}
  \includegraphics[width=0.49\linewidth]{images/52399_mask}
  \includegraphics[width=0.49\linewidth]{images/56885_pred}
  \includegraphics[width=0.49\linewidth]{images/56885_mask}
  \caption{Examples of masks and bounding boxes from detectron (left) and pixel analysis (right).}
  \label{fig:bbox_mask}
\end{figure}

>>>>

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images, e.g. with low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements, provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata,~\cite{biodiv_info_standards} detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images
to determine if, for instance, a specific fish grows to different lengths
in different habitats.
\begin{figure*}
  \centering
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100314}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100682}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100020}
  \caption{Three example images from the Illinois Natural History Survey (INHS) collection.}
  \label{fig:INHS_examples}
\end{figure*}

Since the digital collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible. Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics tools and techniques, we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of 7,247
images drawn from the Illinois Natural History Survey (INHS) Collection.
Figure \ref{fig:INHS_examples} presents three typical images used in our
study.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results are presented, followed by a discussion. The conclusion highlights key findings and identifies next steps.

@@

\section{Methods}
Our process for metadata generation can be broken into three steps: object detection with Facebook's Detectron2 machine learning library (subsequently referred to as \verb|detectron|), image processing at the pixel level, and calculations on the results of the previous steps to determine higher level metadata properties.
% \todo{Definitely need to explain somewhere in more detail about our current admissibility criteria for images. Not sure where I want to put it yet so just leaving it as a todo}

\begin{figure}[b]
  \centering
%  \includegraphics[width=.49\linewidth]{images/teaser1_crop}
  %\hspace{4mm}
  \includegraphics[width=.90\linewidth]{images/teaser2_crop}
  \caption{Initial object detection on a specimen image using Detectron2~\cite{wu2019detectron2}.}
  \label{fig:teaser}
\end{figure}

\subsection{Detectron}
A prerequisite task to performing any advanced metadata property generation
is finding the specimens (and other relevant objects) within the collection
images. Object detection has been a broadly active field of study in recent
years \cite{zou2019object}, and has resulted in a number of well-tested, purpose-built architectures. We elected to use Facebook AI Research's \verb|detectron| tool \cite{wu2019detectron2}, and specifically its implementation of the Mask R-CNN architecture \cite{he2018mask}, for object detection in our project.
\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format datasets. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given image. 
For our project, we use it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers,
as shown in Figure \ref{fig:teaser}.

@@

\begin{table*}[t]
    \centering
    \caption{Metadata properties}
    \label{tab:properties}
    \begin{tabular}{cccp{0.5\linewidth}}
        \toprule
        \textbf{Property} & \textbf{Association} & \textbf{Type} & \textbf{Explanation}\\
        \midrule
        \verb|has_fish| & Overall Image & Boolean & Whether a fish was found in the image.\\
        \verb|fish_count| & Overall Image & Integer & The quantity of fish present.\\
        \verb|has_ruler| & Overall Image & Boolean & Whether a ruler was found in the image.\\
        \verb|ruler_bbox| & Overall Image & 4 Tuple & The bounding box of the ruler (if found).\\
        \verb|scale| & Overall Image & Float & The scale of the image in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|bbox| & Per Fish & 4 Tuple & The top left and bottom right coordinates of the bounding box for a fish.\\
        \verb|background.mean| & Per Fish & Float & The mean intensity of the background within a given fish's bounding box.\\
        \verb|background.std| & Per Fish & Float & The standard deviation of the background within a given fish's bounding box.\\
        \verb|foreground.mean| & Per Fish & Float & The mean intensity of the foreground within a given fish's bounding box.\\
        \verb|foreground.std| & Per Fish & Float & The standard deviation of the foreground within a given fish's bounding box.\\
        \verb|contrast| & Per Fish & Float & The contrast between foreground and background intensities within a given fish's bounding box.\\
        \verb|centroid| & Per Fish & 4 Tuple & The centroid of a given fish's bitmask.\\
        \verb|primary_axis| & Per Fish & 2D Vector & The unit length primary axis (eigenvector) for the bitmask of a given fish.\\
        \verb|clock_value| & Per Fish & Integer & Fish's primary axis converted into an integer ``clock value'' between 1 and 12.\\
        \verb|length| & Per Fish & Float & The length of a fish in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|mask| & Per Fish & 2D Matrix & The bitmask of a fish in 0's and 1's.\\
        \verb|pixel_analysis_failed| & Per Fish & Boolean & Whether the pixel analysis process failed for a given fish. If \verb|true|, \verb|detectron|'s mask and bounding box were used for metadata generation.\\
        \verb|score| & Per Fish & Float & The percent confidence score output by \verb|detectron| for a given fish.\\
        \verb|has_eye| & Per Fish & Boolean & Whether an eye was found for a given fish.\\
        \verb|eye_center| & Per Fish & 2 Tuple & The centroid of a fish's eye.\\
        \verb|side| & Per Fish & String & The side (i.e.\ \verb|'left'| or \verb|'right'|) of the fish that is facing the camera (dependent on finding its eye).\\
      \bottomrule
\end{tabular}
\end{table*}

@@

\subsubsection{contrast}
The contrast between the intensities of the foreground and background
pixels is computed as \verb|background.mean| - \verb|foreground.mean|.

@@

Since we recalculate the mask of the fish, its centroid must be recalculated
as well. This can be done via
\begin{equation}
    (\bar{x}, \bar{y}) = (\mathrm{round}(\frac{M_{10}}{M_{00}}), \mathrm{round}(\frac{M_{01}}{M_{00}})),
\end{equation}
where \(M_{00}\) is the pixel area of the fish's blob, \(M_{10}\) is the sum of all the \(x\) values of blob pixels, and \(M_{01}\) is the sum of all the \(y\) values of blob pixels.
\subsubsection{side}
Determining which \verb|side| of the fish is visible is predicated on finding its eye. If an eye is found, the sign of the \(x\) component of the vector from the centroid of the fish to the centroid of the eye specifies which side is up: negative for left and positive for right. This assumes the fish was photographed vertically, which is essentially always the case for all image collections our group has worked on.
\subsubsection{primary\_axis and clock\_value}
The \verb|primary_axis| of a fish can be calculated by taking the covariance of its blob in \(x\) and \(y\), which yields its principle eigenvector.
The eigenvector can be directly assigned to the property. If an eye is
present, we ensure that \verb|primary_axis| points in the direction of the
eye in relation to the fish's centroid.

@@

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results and Discussion}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{leipzig2021biodiversity} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 50 specimens from the set of \(7,244\) were analyzed by
hand for comparison.

Our automated process currently generates 6 of the 22 core metadata properties: \verb|if_fish| (\verb|has_fish|), \verb|fish_number| (\verb|fish_count|), \verb|if_ruler| (\verb|has_ruler|), \verb|specimen_angled| (\verb|clock_value|), \verb|specimen_view| (\verb|side|), and \verb|brightness| (\verb|foreground.mean|).
% and \verb|if_background_uniform| (\verb|background.std|).
In addition, our process also calculates contrast, bounding boxes and
fish lengths in centimeters.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/none1_crop} \\
  \vspace{3mm}
  \includegraphics[width=0.9\linewidth]{images/double1_crop}
  \caption{A fish that was not detected (top) and a fish that was detected twice (bottom).}
  \label{fig:fish_detect_wrong}
\end{figure}

\subsection{Fish Detection}
All images in the INHS dataset contain exactly one fish. For \(7,209\) of the specimen images, one fish was detected, a 99.5\% correct rate.
For 25 of the images, 2 fish were detected, 3 fish were detected for 3 images, and for 7 images no fish were found.
The 7 fish that were not detected were quite small. This type of specimen 
is currently lacking from the training set.
See the top image in Figure~\ref{fig:fish_detect_wrong} for an example.
In the case of greater than 1 fish, 9 of the 28 contained tags that
overlapped the fish and were themselves labeled as a second fish.
Of the remaining 17, \verb|detectron| erroneously labeled the fish as two separate fish objects, or labeled a subsection of the fish a second time.
Fish that were double labeled were generally quite large and/or dissimilar from the fish found in the training set, such as the bottom image in
Figure~\ref{fig:fish_detect_wrong}.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/no_ruler1}
  \includegraphics[width=0.49\linewidth]{images/no_ruler2}
  \caption{The two images where the ruler was not found. The left image exhibits a yellow hue, and the right is quite washed out with poor contrast.}
  \label{fig:no_ruler}
\end{figure}

\subsection{Ruler Detection}
For all but 2 of the \(7,244\) images \verb|detectron| was able to find the ruler, a nearly perfect correct rate.
In 56 of the images, the ruler itself was found, but the numbers ``2''
and/or the ``3'' on the ruler were not. Therefore a scale calculation could
not be performed, producing a 99.2\% success rate for
the \verb|scale| computation.

Images where one of these objects were not detected generally had some form of coloration issue. They were either washed out, very dark or yellow in hue.
See Figure \ref{fig:no_ruler} two such examples.
Some of the rulers for which ``2'' and/or ``3'' were not detected were particularly scratched and damaged. Only ``3'' was missed in 45 of the 56 cases, only ``2'' was missed in 2 cases, and both were missed in 9 cases. This may indicate that more training samples for ``3'' are required. Many of the rulers where both numerals were missed were particularly small within the image, which again may be solvable through expanding the training dataset.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.31\linewidth]{images/dark}
  \includegraphics[width=0.31\linewidth]{images/normal}
  \includegraphics[width=0.31\linewidth]{images/bright}
  \caption{Examples of low, medium and high contrast fish images.}
  \label{fig:contrast}
\end{figure*}

@@

% There were six cases where an eye was detected but the \verb|side| value
% was wrong.
For these 6 cases, a spot on the wrong side of the fish was labeled as the
most likely eye within the bounding box of the fish, i.e.~eye detection 
was incorrect. Figure \ref{fig:wrong_eye} presents one such example.
There were an additional 17 images for which the automated process generated a result that did not match the manually created data. For these remaining cases the manual data was incorrect, giving the automated system an error rate \(2.8\) times lower than the human error rate.
This result highlights the additional utility of automated methods to
double-check and verify human-generated metadata.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.49\linewidth]{images/wrong_side_orig}
  \includegraphics[width=0.49\linewidth]{images/wrong_side1}
  \caption{A fish for which a splotch on its tail fin was labeled the most likely eye.}
  \label{fig:wrong_eye}
\end{figure}

\subsection{Clock Value}
Clock position values were successfully generated for \(6,991\) of the images. Of those, all but 8 were within \(\pm{}1\) of the correct result,
our definition of a correct/acceptable result, making the correct rate for
this computation 96.4\%.

Of the specimens for which clock values were generated, 33 did not match the manually created data (within a tolerance of \(\pm{}1\)).
For 25 of those, the human-generated data was incorrect, giving the automated process a \(3.1\) times lower error rate. Of the 8 that were
computationally classified incorrectly,
two specimens were quite curved making it difficult to assign a clear angle value, as
seen in Figure \ref{fig:curvedFish}.
The other 6 were values of ``3'' instead of ``9'' or vice versa, which resulted from
a mislabeled eye as discussed in the previous section.
%\todo{Why is the following sentence here? What does it have to do with
%Clock Value?}
%A small number of images (such as on the right in Figure~\ref{fig:teaser}) have tags that overlap with the fish. These tags were light enough that they were correctly labeled as background by the pixel analysis.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\linewidth]{images/curved1}
  \caption{An example of a heavily curved specimen.}
  \label{fig:curvedFish}
\end{figure}

\begin{comment}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/histo}
  \caption{Histogram for difference in brightness between foreground an background.}
  \label{fig:histo}
\end{figure}
\end{comment}

\subsection{Image Contrast}
The contrast of an image is an important image property needed
for the analyses of the BGNN study. Ideally, for INHS images the fish should be well lit and clearly displayed, and the background should also be as light and white as possible.
% Figure~\ref{fig:histo} shows the distribution of
The difference between
the mean intensity of the foreground (i.e.\ the pixels of the fish) and the mean intensity of the background within the fish's bounding box
was computed for all 7,244 images.
The overall mean of the differences is \(144.3\), with a standard deviation of \(15.8\). Images on the low end of the distribution exhibit poor foreground--background contrast, and images on the high end likely contain poorly lit specimens.
Images are considered to have ``low'' and
``high'' contrast if their background-foreground difference is greater than
one standard deviation away from the mean; otherwise they are classified as
``medium''.
Examples of each type of image (low, medium, and high contrast) can be seen in Figure~\ref{fig:contrast}.
The left image has a mean intensity of 103 (low), the middle image 144
(medium), and the right 186 (high).

\begin{table}[H]
    \centering
%    \caption{\verb|foreground.mean| statistics for the three Brightness classes}
    \caption{foreground.mean statistics for the three Brightness classes}
    \label{tab:bright}
    \begin{tabular}{cccc}
        \toprule
        \textbf{Class} & \textbf{Mean Intensity} & \textbf{Standard Deviation} & \textbf{Instances}\\
        \midrule
        Dark & 75.2 & 13.5 & 1800\\
        Normal & 93.6 & 14.8 & 5186\\
        Bright & 108.4 & 15.2 & 242\\
        \bottomrule
    \end{tabular}
\end{table}
\subsection{Brightness}
Specimen brightness is one of the 22 hand-recorded metadata properties. It is encoded as \verb|dark|, \verb|normal| or \verb|bright|. These values
correspond to the mean foreground intensity computed by the automated system.
The mean and standard deviation of the mean foreground intensities were
computed for the images in the three human-specified classes.
Table~\ref{tab:bright} contains the resulting values and shows that
automated intensities values provide objective measures that may be used to break images into groups that correlate with human-defined brightness
classifications.
The mean of \verb|foreground.mean| over all 7,244 images is 89.5, with a
standard deviation of 16.9. These numbers suggest, for example, that
\verb|foreground.mean| $\pm$ one-half a standard deviation could define 
\verb|normal| brightness, with values outside this range being classified
as \verb|dark| or \verb|bright|.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.49\linewidth]{images/011196_pred}
  \includegraphics[width=0.49\linewidth]{images/011196_mask}
  \includegraphics[width=0.49\linewidth]{images/52399_pred}
  \includegraphics[width=0.49\linewidth]{images/52399_mask}
  \includegraphics[width=0.49\linewidth]{images/56885_pred}
  \includegraphics[width=0.49\linewidth]{images/56885_mask}
  \caption{Examples of masks and bounding boxes from detectron (left) and pixel analysis (right).}
  \label{fig:bbox_mask}
\end{figure}

Fish bounding boxes were calculated for all \(7,237\) images in which a fish was found. All but 263 of these were generated via pixel analysis, with those 263 falling back to the original \verb|detectron| bounding box.
50 randomly-chosen images were reviewed manually to evaluate the
calculation, a representative sample are presented in
Figure \ref{fig:bbox_mask}.
All 50 masks and bounding boxes were correctly placed
on/around the location of the fish. However, a number of them lacked portions of lightly colored tails and/or fins.
Specifically, 22 masks and bounding boxes covered the entire fish,
16 missed some of the tail, and 12 missed most or all of the tail, as
seen in Figure \ref{fig:missedTail}.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/61631}
  \includegraphics[width=0.49\linewidth]{images/61631_mask}
  \caption{An example of a light colored tail being missed during the pixel analysis process.}
  \label{fig:missedTail}
\end{figure}
Masks and bounding boxes contain the head and trunk of the fish in nearly all cases, but further refinement of our algoriths will be needed
to ensure that light fins and tails are masked consistently and accurately.
